---
title: "PMR"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pander)
library(kableExtra)
library(cowplot) #for nice plot plot_grid

#manage function of leaf/flwering phenology 
functions <- list.files("./pheno_functions", full.names = T) %>%
  map(source)

```

## General description

functions to simulate budburst with R, translation of the PMP plateforme (Chuine et al 2013)

## Data information

data used for the illustration are coming from Mt-Ventoux (contact: hendrik.davi@inrae.fr)

### Climate data
data along elevation...

```{r data climate presentation, echo = FALSE,results='hide',fig.keep='all', message=FALSE}

pathClimate <- "/home/journe/Documents/DATA/LOCAL_CLIMATE/Ventoux/"
files_climate <- list.files(pathClimate, full.names = F, pattern = "*.txt") #, recursive = T

#based on code from C. Wilke
VtxClimate <- tibble(filename = files_climate) %>% # create a data frame
  # holding the file names
  mutate(file_contents = map(filename,          # read files into
                             ~ read_delim(file.path(pathClimate, .), col_names = TRUE, delim = "\t", skip = 12, escape_double = FALSE, trim_ws = TRUE))) %>%
  unnest(cols = c(file_contents)) %>% 
  select(-gr, -ws) %>% #remove global radiation and wind speed 
  rename(plot= filename, year=`# y`, month=m, DOY=d, humidityRelative = rh, precipitation = p,maxTemp=tmax, minTemp = tmin, meanTemp=tj) %>% 
  mutate(plot = str_extract(plot, "[^_]+")) %>% 
  filter(grepl("DVX", plot))

VtxClimate %>% group_by(plot, year) %>% 
  summarise(meanT = mean(meanTemp), sdT = sd(meanTemp)) %>% 
ggplot(aes(x = year, y = meanT, col = plot))+geom_point()+geom_line()+xlab("Year")+ylab("Mean temperature")

```

### Phenology data
leaf budburst from Mt Ventoux


## Model information

### Climate data format

- ```plot```: 1 plot for each use of the function. If you have many more plots use a loop
- ```year```: numbers of years. Not that the function need at least two year to run, because it use bot climate of the previous $n-1$ and current year $n$ to simulate leaf budburst of the year $n$.
- ```DOY```: day of the year, in the following format 1 to X, where if you have for example two years, X should go to 730 days. The current function can not work with DOY going to 365 days only. 
- ```meanTemp```: mean temperature for each days in T°C. 
For each function you need to specify date of the origin of your climate file. 

### Phenological models 
#### Unichill model
In the current version, unichill is based on two different options. A threshold model to accumate temperature of rate of chilling or exponential model as described in Chuine (2000). In a recent article Gauzere et al (2020) found that response of chilling was simulated well with a sigmoïd function rather than a threshold model. So if you plan to use this function you need to specify a new argument (deciduous or coniferous). The sigmoid submodel need more parameters than the threshold submodel. More details are provided in Chuine (2000) and in SI from Gauzere et al (2020). 

Parameters are in the following order, for example if it is a coniferous (here example with Abies parameters) : $t$; $t0$; $F_crit$; $T50$; $dT$; $C_crit$; $a$; $b$; $c$. The last parameters are parameters for the sigmoid function of rate of chilling.
```parametersAbies = c(-67, 9, 15.2, 0.24, 144.6, 1.61, -27.1, -4.95)```

Parameters for threshold option, for example here for Fagus are in the following order: $t0$; $F_crit$; $T50$; $dT$; $C_crit$; $T_base$. In this submodel you need to specify a base temperature. 
```parametersFagus = c(-62, 37.3, 13, 0.08, 73.7, 13)```

We first test here unichill model for \textit{Fagus sylvatica}
```{r fitUnichill fagus,results='hide'}
#select data for the site of interest (here named DVX5)
dataDvx5= VtxClimate %>% filter(plot == "DVX5")
#need to specify origin data of the data in this particular format
originClimateData <- "1959-01-01"
#need to specify plant functional type. 
FunctionalTypeFagus <- "deciduous"

#species parameter here for Fagus sylvatica
parametersFagus = c(-62, 37.3, 13, 0.08, 73.7, 13)

#Run unichill model in plot DVX5 for fagus and then add plot and species name 
phenoFagusDvx5 <-   Unichill_Threshold_Chuine(parametersFagus, dataDvx5, originClimateData, FunctionalTypeFagus) %>% 
  mutate(species = "Fagus", site = "DVX5")

```

Then we could also test for \textit{Abies alba} with the same previous climate (DVX5) and same climate origin. 

```{r fitUnichill abies,results='hide'}
#need to specify plant functional type. 
FunctionalTypeAbies <- "coniferous"

#species parameter here for Abies alba
parametersAbies = c(-67, 9, 15.2, 0.24, 144.6, 1.61, -27.1, -4.95)

#Run unichill model in plot DVX5 for Abies and add details with dplyr
phenoAbiesDvx5 <-   Unichill_Threshold_Chuine(parametersAbies, dataDvx5, originClimateData, FunctionalTypeAbies) %>% 
  mutate(species = "Abies", site = "DVX5")

```

You can see a model output in the following figure of leaf bud burst simulate for the two species. 
```{r sim unichill, echo = FALSE,results='hide',fig.keep='all', message=FALSE}

rbind(phenoAbiesDvx5, phenoFagusDvx5) %>% 
  drop_na() %>% 
  ggplot(aes(x = year, y = BBDAY, col = species))+geom_point()+geom_line()



```



## References 
- Chuine, I. 2000. A unified model for budburst of trees. J. Theor. Biol. 207: 337– 347. 
- Chuine I., Garcia de Cortazar Atauri I., Kramer K. & Hänninen H. (2013) Plant Development Models. In: Phenology: An Integrative Environmental Science (ed. Schwarz MD). Springer, Dordrecht, Netherlands, pp. 275-293.
- Gauzere J, Teuf B, Davi H, Chevin L.‐M., Caignard T, Delzon S, Ronce O, Chuine I. 2020. Where is the optimum? Predicting the variation of selection along climatic gradients and the adaptive value of plasticity. A case study on tree phenology. Evolution Letters 4: 109– 123. 