---
title: "PMR"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(pander)
library(kableExtra)
library(cowplot) #for nice plot plot_grid

#manage function of leaf/flwering phenology 
functions <- list.files("./pheno_functions", full.names = T) %>%
  map(source)

```

## General description

functions to simulate budburst with R, translation of the PMP plateforme (Chuine et al 2013)

## Data information

data used for the illustration are coming from Mt-Ventoux (contact: hendrik.davi@inrae.fr)

### Climate data
data along elevation...

```{r data climate presentation, echo = FALSE,results='hide',fig.keep='all', message=FALSE}

pathClimate <- "/home/journe/Documents/DATA/LOCAL_CLIMATE/Ventoux/"
files_climate <- list.files(pathClimate, full.names = F, pattern = "*.txt") #, recursive = T

#based on code from C. Wilke
VtxClimate <- tibble(filename = files_climate) %>% # create a data frame
  # holding the file names
  mutate(file_contents = map(filename,          # read files into
                             ~ read_delim(file.path(pathClimate, .), col_names = TRUE, delim = "\t", skip = 12, escape_double = FALSE, trim_ws = TRUE))) %>%
  unnest(cols = c(file_contents)) %>% 
  select(-gr, -ws) %>% #remove global radiation and wind speed 
  rename(plot= filename, year=`# y`, month=m, DOY=d, humidityRelative = rh, precipitation = p,maxTemp=tmax, minTemp = tmin, meanTemp=tj) %>% 
  mutate(plot = str_extract(plot, "[^_]+")) %>% 
  filter(grepl("DVX", plot))

VtxClimate %>% group_by(plot, year) %>% 
  summarise(meanT = mean(meanTemp), sdT = sd(meanTemp)) %>% 
ggplot(aes(x = year, y = meanT, col = plot))+geom_point()+geom_line()+xlab("Year")+ylab("Mean temperature")

```

### Phenology data
leaf budburst from Mt Ventoux


## Model information

### Climate data format

- ```plot```: 1 plot for each use of the function. If you have many more plots use a loop
- ```year```: numbers of years. Not that the function need at least two year to run, because it use bot climate of the previous $n-1$ and current year $n$ to simulate leaf budburst of the year $n$.
- ```DOY```: day of the year, in the following format 1 to X, where if you have for example two years, X should go to 730 days. The current function can not work with DOY going to 365 days only. *** works also with basic DOY from 1 to 365*** 
- ```meanTemp```: mean temperature for each days in T°C. 
For each function you need to specify date of the origin of your climate file. 

### Phenological models 
#### Unichill model
In the current version, unichill is based on two different options. A threshold model to accumulate temperature of rate of chilling or exponential model as described in Chuine (2000). In a recent article Gauzere et al (2020) found that response of chilling was simulated well with a sigmoïd function rather than a threshold model. So if you plan to use this function you need to specify a new argument (unimodal or threshold). The sigmoid submodel need more parameters than the threshold submodel. More details are provided in Chuine (2000) and in SI from Gauzere et al (2020). 

Parameters are in the following order, for example if it is a unimodal submodel (here example with Abies parameters) : $t$ ; $t0$ ; $F_{crit}$ ; $T50$ ; $dT$ ; $C_{crit}$ ; $a$ ; $b$ ; $c$ . The last parameters are parameters for the sigmoid function of rate of chilling.
```parametersAbies = c(-67, 9, 15.2, 0.24, 144.6, 1.61, -27.1, -4.95)```

Parameters for threshold option, for example here for Fagus are in the following order: $t0$ ; $F_{crit}$ ; $T50$ ; $dT$ ; $C_{crit}$ ; $T_{base}$ . In this submodel you need to specify a base temperature. 
```parametersFagus = c(-62, 37.3, 13, 0.08, 73.7, 13)```

We first test here unichill model for \textit{Fagus sylvatica}
```{r fitUnichill fagus,results='hide'}
#select data for the site of interest (here named DVX5)
dataDvx5= VtxClimate %>% filter(plot == "DVX5")

#need to specify submodel of unichill 
submodelFagus <- "threshold"

#species parameter here for Fagus sylvatica with the folowing order
#t0, Fcrit, T50, Dt, Ccrit, Tbase
parametersFagus = c(-62, 37.3, 13, 0.08, 73.7, 13)

#Run unichill model in plot DVX5 for fagus and then add plot and species name 
phenoFagusDvx5 <-   Unichill_Chuine(parametersFagus, dataDvx5, submodelFagus) %>% 
  mutate(species = "Fagus", site = "DVX5", model = "Unichillth")

```

Then we could also test for \textit{Abies alba} with the same previous climate (DVX5) and same climate origin. 

```{r fitUnichill abies,results='hide'}
#need to specify submodel of unichill
submodelAbies <- "unimodal"

#species parameter here for Abies alba
#t0, Fcrit, T50, Dt, Ccrit, Tbase, a , b , c
parametersAbies = c(-67, 9, 15.2, 0.24, 144.6, 1.61, -27.1, -4.95)

#Run unichill model in plot DVX5 for Abies and add details with dplyr
phenoAbiesDvx5 <-   Unichill_Chuine(parametersAbies, dataDvx5, submodelAbies) %>% 
  mutate(species = "Abies", site = "DVX5", model = "Unichill")

```

You can see a model output in the following figure of leaf bud burst simulate for the two species. 
```{r sim unichill, echo = FALSE,results='hide',fig.keep='all', message=FALSE}

rbind(phenoAbiesDvx5, phenoFagusDvx5) %>% 
  drop_na() %>% 
  ggplot(aes(x = year, y = BBDAY, col = species))+geom_point()+geom_line()



```

#### Uniforc model
The uniforc model is a simple model where temperature are accumlated the same year of leaf budburst. It has four parameters, already define previously: $t0$ ; $F_{crit}$ ; $T50$ ; $dT$ . Here, $t0$ is the starting date of ecodormancy and not endodormancy as in unichill model. Parameters defined here came from Gauzere et al (2017)

```{r fitUniforc fagus,results='hide'}
#select data for the site of interest (here named DVX5)
dataDvx5= VtxClimate %>% filter(plot == "DVX5")

#species parameter here for Fagus sylvatica
parametersFagusUniforc = c(13, 41.3, 12.41, 0.06)
#order : t0, Fcrit, T50, dT
#Run unichill model in plot DVX5 for fagus and then add plot and species name 
phenoFagusDvx5Unif <-   Uniforc_Chuine(parametersFagusUniforc, dataDvx5) %>% 
  mutate(species = "Fagus", site = "DVX5", model = "Unifor")

```

```{r sim uniforc, echo = FALSE,results='hide',fig.keep='all', message=FALSE}

phenoFagusDvx5Unif%>% 
  drop_na() %>% 
  ggplot(aes(x = year, y = BBDAY))+geom_point()+geom_line()

```

#### PGC model - Gauzere model
The PGC model is a model with temperature and photoperiod effect to simulate budburst. This model showed robust prediction of leaf budburst of Fagus sylvatica. However here we need to add a new column with the latitude to calculated daylength (i.e. photoperiod). The model has 9 parameters : $t0$ ; $F_{crit}$ ; $T50$ ; $dT$ ; $T_{base}$ ; $C50$ ; $pR$ ; $dC$ ; $dP$ 


```{r PGC fagus,results='hide'}
#select data for the site of interest (here named DVX5)
dataDvx5= VtxClimate %>% filter(plot == "DVX5")

#species parameter here for Fagus sylvatica
parametersFagusPGC = c(-65, 13.9, 10.5, 0.25, 11.9, 153.5, 1.76, -40, 0.44)
latitudeDvx5 = 44.18 #in degree
#order : $t0$ ; $F_{crit}$ ; $T50$ ; $dT$ ; $T_{base}$ ; $C50$ ; $pR$ ; $dC$ ; $dP$
    # t0 <- parameters[1] # starting date of ecodorimancy
    # F_crit <- parameters[2] # Critical state of forcing
    # T50 <- parameters[3] # Mid-response temperature to forcing
    # dT <- parameters[4] # Slope of the forcing response
    # T_base <- parameters[5] # Threshold chilling temperature
    # C50 <- parameters[6] # Mid-response photoperiod
    # pR <- parameters[7] #boudaries mid repsonse photoperiod
    # dC <- parameters[8] #slope photoperiod
    # dP <- parameters[9] #slope growht competence
#Run unichill model in plot DVX5 for fagus and then add plot and species name 
phenoFagusDvx5PGC <-   PGC(parametersFagusPGC, dataDvx5, latitudeDvx5) %>% 
  mutate(species = "Fagus", site = "DVX5", model = "PGC")

```

```{r sim PGC, echo = FALSE,results='hide',fig.keep='all', message=FALSE}

phenoFagusDvx5PGC%>% 
  drop_na() %>% 
  ggplot(aes(x = year, y = BBDAY))+geom_point()+geom_line()

```

#### Comparison between the different models


```{r sim comparison, echo = FALSE,results='hide',fig.keep='all', message=FALSE}
#comparison of the three model for fagus 

dfcomp <- rbind(phenoFagusDvx5, phenoFagusDvx5Unif, phenoFagusDvx5PGC)


dfcomp%>% 
  drop_na() %>% 
  ggplot(aes(x = year, y = BBDAY, col = model))+geom_point()+geom_line()

dfcomp%>% 
  drop_na() %>% 
  ggplot(aes(x = model, y = BBDAY, col = model))+geom_boxplot()


```

## References 
- Chuine, I. 2000. A unified model for budburst of trees. J. Theor. Biol. 207: 337– 347. 
- Chuine I., Garcia de Cortazar Atauri I., Kramer K. & Hänninen H. (2013) Plant Development Models. In: Phenology: An Integrative Environmental Science (ed. Schwarz MD). Springer, Dordrecht, Netherlands, pp. 275-293.
- Gauzere J, Teuf B, Davi H, Chevin L.‐M., Caignard T, Delzon S, Ronce O, Chuine I. 2020. Where is the optimum? Predicting the variation of selection along climatic gradients and the adaptive value of plasticity. A case study on tree phenology. Evolution Letters 4: 109– 123. 
- J. Gauzere, S. Delzon, H. Davi, M. Bonhomme, A. Garcia de Cortazar, I. Chuine. Integrating interactive effects of chilling and photoperiod in phenological process-based models. A case study with two European tree species: Fagus sylvatica and Quercus petraea
Agric. Forest Meteorol., 244 (2017), pp. 9-20